---
title: "Regional Analysis of Global Displacement"
author: "Chao Huang"
date: 'Version: `r Sys.Date()`'
output:
  pagedown::html_paged:
    toc: true
    # change to true for a self-contained document, but it'll be a litte slower for Pandoc to render
    self_contained: true
    highlight: null
    toc-title: Contents
    paged-footnotes: true
    css:
    - data/css/unhcr-page.css
    - data/css/unhcr.css
    - data/css/unhcr-fonts.css 
---

# Executive summary

In the last decade, The global population of forcibly displaced people grew substantially from 43.3 million in 2009 to 70.8 million in 2018, reaching a record high. Now UNHCR is standing at the start point of the new decade, and it is worthwhile to establish a predictive model for strategic planning. Instead of predicting the absolute value of global displacement, we are looking at the concept of 'Displace Rate', which is the ratio of the displaced population and the total population. Moreover, the data is disaggregated by the UNHCR regional bureaus. Apart from temporal analysis, we also build a dynamic regression model to explore the connection between displacement and conflicts. This implies 3 distinct steps presented in the chapters report:

1. Consolidate the data into time series;
2. Apply time series analysis and dynamic regression model with the data of conflicts;
3. Model diagnosis.

The code from this report is fully reproducible and based on public data sources. It can be regenerated automatically with newly released data, and all prediction will be adjusted accordingly.



```{r global_options, echo=FALSE, message=FALSE, include=FALSE}
knitr::opts_chunk$set(error = FALSE, warning = FALSE, cache = TRUE, message = FALSE)

options(scipen = 999) # turn-off scientific notation like 1e+48

## Getting all necessary package
using <- function(...) {
    libs <- unlist(list(...))
    req <- unlist(lapply(libs,require,character.only = TRUE))
    need <- libs[req == FALSE]
    if (length(need) > 0) { 
        install.packages(need)
        lapply(need,require,character.only = TRUE)
    }
}
using(
## Data      
      'gapminder', 
      'wbstats',
      'readxl',
      'haven',
## Mapping      
      'ISOcodes',
      'sf',
      'rnaturalearth',
      'rnaturalearthdata',
      'ggspatial',
#plots
      'R.utils',
      'png', 
      'grid',
      'ggplot2',
      'gganimate',
      'gghighlight',
      'ggpubr', 
      'ggalt', 
      'plotly',
      'reshape2',
      'scales', # Scale Functions for Visualization
# manipulation'tidyverse',
      'dplyr', 
      'tidyr',
      'zoo',
      'xml2',
      'tidyverse',
      'forcats', # Tools for Working with Categorical Variables 
#time series
      'dynlm', #Dynamic Linear Regression
      'dynpanel', # Dynamic Panel Data Models
      'plm', # Linear Models for Panel Data
      'urca', # Unit Root and Cointegration Tests for Time Series Data
      'tseries', # Time Series Analysis and Computational Finance
      'TSA', # Time Series Analysis
# modeling
      'car', # Companion to Applied Regression
      'msm', # Multi-State Markov and Hidden Markov Models in Continuous Time 
      'pscl', # Political Science Computational Laboratory 
      'boot', # Bootstrap Functions
#forecasting
      'forecast',
      'prediction',
      'estimatr',
      'rgeos',
#knitting
      'markdown',
      'pander',
      'knitr',
      'rvest',
      'stargazer',
      'texreg',
      'ggfortify'
      )

unhcr_style <- function() {
  font <- "Lato"
  ggplot2::theme(
    
#This sets the font, size, type and colour of text for the chart's title
  plot.title = ggplot2::element_text(family = font, size = 20, face = "bold", color = "#222222"),

#This sets the font, size, type and colour of text for the chart's subtitle,  as well as setting a margin between the title and the subtitle
  plot.subtitle = ggplot2::element_text(family = font, size = 16, margin = ggplot2::margin(9,0,9,0)),
  plot.caption = ggplot2::element_blank(),

#This sets the position and alignment of the legend, removes a title and backround for it and sets the requirements for any text within the legend. The legend may often need some more manual tweaking when it comes to its exact position based on the plot coordinates.
  legend.position = "top",
  legend.text.align = 0,
  legend.background = ggplot2::element_blank(),
  legend.title = ggplot2::element_blank(),
  legend.key = ggplot2::element_blank(),
  legend.text = ggplot2::element_text(family = font, size = 13, color = "#222222"),

#This sets the text font, size and colour for the axis test, as well as setting the margins and removes lines and ticks. In some cases, axis lines and axis ticks are things we would want to have in the chart
  axis.title = ggplot2::element_blank(),
  axis.text = ggplot2::element_text(family = font, size = 13, color = "#222222"),
  axis.text.x = ggplot2::element_text(margin = ggplot2::margin(5, b = 10)),
  axis.ticks = ggplot2::element_blank(),
  axis.line = ggplot2::element_blank(),

#This removes all minor gridlines and adds major y gridlines. In many cases you will want to change this to remove y gridlines and add x gridlines. 
  panel.grid.minor = ggplot2::element_blank(),
  panel.grid.major.y = ggplot2::element_line(color = "#cbcbcb"),
  panel.grid.major.x = ggplot2::element_blank(),

#This sets the panel background as blank, removing the standard grey ggplot background colour from the plot
  panel.background = ggplot2::element_blank(),

#This sets the panel background for facet-wrapped plots to white, removing the standard grey ggplot background colour and sets the title size of the facet-wrap title to font size 22
  strip.background = ggplot2::element_rect(fill = "white"),
  strip.text = ggplot2::element_text(size  = 13,  hjust = 0)
  )
}

#Left align text
left_align <- function(plot_name, pieces){
  grob <- ggplot2::ggplotGrob(plot_name)
  n <- length(pieces)
  grob$layout$l[grob$layout$name %in% pieces] <- 2
  return(grob)
}


## a little help function to better format numbers
format_si <- function(...) {
  function(x) {
    limits <- c(1e-24, 1e-21, 1e-18, 1e-15, 1e-12,
                1e-9,  1e-6,  1e-3,  1e0,   1e3,
                1e6,   1e9,   1e12,  1e15,  1e18,
                1e21,  1e24)
    prefix <- c("y",   "z",   "a",   "f",   "p",
                "n",   "",   "m",   " ",   "k",
                "M",   "G",   "T",   "P",   "E",
                "Z",   "Y")

    # Vector with array indices according to position in intervals
    i <- findInterval(abs(x), limits)

    # Set prefix to " " for very small values < 1e-24
    i <- ifelse(i == 0, which(limits == 1e0), i)

    paste(format(round(x/limits[i], 1),
                 trim = TRUE, scientific = FALSE, ...),
          prefix[i])
  }
}


```
# Dataset used for Analysis 
## Regional displaced rate across years 
The displacement data are from UNHCR and disaggregated with UNHCR regional bureaus, which are Regional Bureau for Asia and the Pacific, Europe, the Americas, the Middle East and North Africa, West and Central Africa, the East and Horn of Africa (and Great Lakes) and South Africa.


```{r results='hide', warning=FALSE, message=FALSE, echo=FALSE}


regiondata <- read.csv("data/regioandata.csv")
regiondata <- regiondata[,-1]

regiondata <- regiondata %>% dplyr::select(year,iso,bureau,gd,population,gdp,death)%>% 
  na.omit() %>% group_by(year,bureau) %>% 
  summarise(pop=sum(population),gdp=sum(gdp),gd=sum(gd),death=sum(death))

region.dpr <- regiondata %>% mutate(rdpr = gd*1000/pop,ldeath=log(death))
region.dpr$ldeath[region.dpr$death==0] <- 0
region.dpr$bureau[region.dpr$bureau=="Regional Bureau for Asia and the Pacific"] <- "Asia and Pacific"
region.dpr$bureau[region.dpr$bureau=="Regional Bureau for Europe"] <- "Europe"
region.dpr$bureau[region.dpr$bureau=="Regional Bureau for the Americas"] <- "Americas"
region.dpr$bureau[region.dpr$bureau=="Regional Bureau for the East and Horn of Africa, and Great Lakes"] <- "East and Horn of Africa"
region.dpr$bureau[region.dpr$bureau=="Regional Bureau for the Middle East and North Africa"] <- "MENA"
region.dpr$bureau[region.dpr$bureau=="Regional Bureau for Southern Africa"] <- "Southern Africa"
region.dpr$bureau[region.dpr$bureau=="Regional Bureaufor West and Central Africa"] <- "West and Central Africa"

plot1 <- ggplot(region.dpr,aes(x=year,y=gd))+
  geom_line()+
  scale_color_manual( values = "0072BC") +
  facet_wrap(~bureau, ncol = 4)+
  coord_cartesian(ylim = c(11000, 20000000), xlim = c(1993, 2018)) +
#  scale_y_continuous(label = format_si()) + ## Format axis number
#  unhcr_style() + 
  labs(title = "Displaced people (Absolute value) in different regions",
       subtitle = "(1993 - 2018)",
       x = "Year", 
       y = "Displaced People") 
ggpubr::ggarrange(left_align(plot1, c("subtitle", "title")), ncol = 1, nrow = 1)  

```

In order to smooth the data, we divided the displaced people with the total population in the region and transformed the data into a ratio. The population data are from World Poplation Prospects.
```{r results='hide', warning=FALSE, message=FALSE, echo=FALSE}

plot2 <- ggplot(region.dpr,aes(x=year,y=rdpr))+
  geom_line()+
  facet_wrap(~bureau, ncol = 4)+
    scale_color_manual( values = "0072BC") +
  coord_cartesian(ylim = c(0, 60), xlim = c(1993, 2018)) +
  #  scale_y_continuous(label = format_si()) + ## Format axis number
  #  unhcr_style() + 
  labs(title = "Displaced rate in different regions",
       subtitle = "(1993 - 2018)",
       x = "Year", 
       y = "Displaced People") 
ggpubr::ggarrange(plot2, ncol = 1, nrow = 1,align = "hv")  

```

## Displacement and conflicts
The global displacement is mainly due to persecution, conflict, violence, or human rights violations, i.e. most of this increase was between 2012 and 2015, driven mostly by the Syrian conflict. Apparently, the conflict cannot be ignored in the analysis. We included the Uppsala conflict data within the dataset, and tried to reveal the connection between displaced rate and deaths of civillians in conflicts.

```{r results='hide', warning=FALSE, message=FALSE, echo=FALSE}
ggplot(region.dpr)+
  geom_line(aes(x=year,y=rdpr,group=1,color="Displaced Rate"))+
  geom_line(aes(x=year,y=ldeath,group=2,color="Log of Death in Conflict"))+
  scale_shape_manual(values=c(4,5)) + 
  scale_color_manual( values = c("#338EC9","#F5C205")) +
  facet_wrap(~bureau, ncol = 4)+
  coord_cartesian(ylim = c(0, 60), xlim = c(1993, 2018)) +
  theme(legend.position="bottom")+
  #  scale_y_continuous(label = format_si()) + ## Format axis number
  #  unhcr_style() + 
  labs(title = "Displacement and conflicts in different regions",
       subtitle = "(1993 - 2018)",
       x = "Year", 
       y = "Displaced People") 
```

# Regional Model For Asia and the Pacific

## Dynamic Regression Models

From the curve we could find the conflict data aligns with the displacement in Asia and Pacific area. We then looked into this specific area.

```{r results='hide', warning=FALSE, message=FALSE, echo=FALSE}

AsiaP <- region.dpr %>% filter(bureau=="Asia and Pacific") %>% 
  dplyr::select(rdpr,ldeath) %>% ts(start = c(1993,1),
                                    frequency = 1)
autoplot(AsiaP[,2:3], facets=TRUE) +
  xlab("Year") + ylab("") +
  ggtitle("Displaced Rate in Asia and Pacific")

```

The data are clearly already stationary (as we are considering percentage changes rather than raw displacement), so there is no need for any differencing. The fitted model is

$y_{t} = 0.1878 x_{t}+\eta_{t}$

$\eta_{t}=0.7464 \eta_{t-1}-0.5417 t_{t-2}+\varepsilon_{t}$

$\varepsilon_{t} \sim N(0,0.322)$

Whose details are displaced below, with ARIMA parameters ARIMA(2,0,0).

```{r warning=FALSE, message=FALSE, echo=FALSE}

fit.AsiaP <- auto.arima(AsiaP[,2],xreg=AsiaP[,3])
summary(fit.AsiaP)

```

## Model Evaluation

We evaluated the model by checking the Residual from regression with ARIMA

```{r warning=FALSE, message=FALSE, echo=FALSE}

checkresiduals(fit.AsiaP)

```
Ljung Box Test shows the model is lack of fit to some extent, which might appear due to the insufficient amount of data. Meanwhile, the histogram of the residuals shows one positive outlier, which will also affect the coverage of the prediction intervals.

## Forecasting

To forecast using a regression model with ARIMA errors, we need to forecast the regression part of the model and the ARIMA part of the model, and combine the results. As with ordinary regression models, in order to obtain forecasts we first need to forecast the predictors. When the predictors are known into the future (e.g., calendar-related variables such as time, day-of-week, etc.), this is straightforward. But when the predictors are themselves unknown, we must either model them separately, or use assumed future values for each predictor.

We will calculate forecasts for the next decade assuming that the future deaths in conflicts will be equal to the mean death from the last 20 years.


```{r results='hide', warning=FALSE, message=FALSE, echo=FALSE}

fcast.Asiap <- forecast(fit.AsiaP,xreg = rep(mean(AsiaP[,3]),12))
autoplot(fcast.Asiap) + xlab("Year") +
  ylab("Displaced Rate")

```

We also compared the ture value nad fitted value, as shown below,

```{r results='hide',warning=FALSE, message=FALSE, echo=FALSE}

plot.fitted <- autoplot(AsiaP[,2], series="True Value") +
  autolayer(fitted(fit.AsiaP), series="Fitted Value") +
  ylab("") +
  ggtitle("Displaced Rate in Asia and the Pacific Area") +
  guides(colour=guide_legend(title=" "))+
  theme(text = element_text(family = "STHeiti"))+
  theme(plot.title = element_text(hjust = 0.5))
ggpubr::ggarrange(left_align(plot.fitted, c("subtitle", "title")), ncol = 1, nrow = 1) 
```






